\documentclass[12pt]{amsart}
%prepared in AMSLaTeX, under LaTeX2e
\addtolength{\oddsidemargin}{-.65in} 
\addtolength{\evensidemargin}{-.65in}
\addtolength{\topmargin}{-.4in}
\addtolength{\textwidth}{1.4in}
\addtolength{\textheight}{0.8in}

\renewcommand{\baselinestretch}{1.05}

\usepackage{xspace,bm}
\usepackage{verbatim,fancyvrb}
\usepackage{palatino}
\usepackage[final]{graphicx}
\usepackage[pdftex,colorlinks=True]{hyperref}

\newtheorem*{thm}{Theorem}
\newtheorem*{defn}{Definition}
\newtheorem*{example}{Example}
\newtheorem*{problem}{Problem}
\newtheorem*{remark}{Remark}

\newcommand{\mfile}[1]{
\begin{quote}
\bigskip
\VerbatimInput[frame=single,framesep=3mm,label=\fbox{\normalsize \textsl{\,#1\,}},fontfamily=courier,fontsize=\footnotesize]{../matlab/#1}
\medskip
\end{quote}
}

%\DefineVerbatimEnvironment{mVerb}{Verbatim}{numbersep=2mm,frame=lines,framerule=0.1mm,framesep=2mm,xleftmargin=4mm,fontsize=\footnotesize}
\DefineVerbatimEnvironment{mVerb}{Verbatim}{numbersep=2mm,xleftmargin=4mm,fontsize=\footnotesize}

% macros
\usepackage{amssymb}

\newcommand{\ba}{\bm{a}}
\newcommand{\bb}{\bm{b}}
\newcommand{\bc}{\bm{c}}
\newcommand{\br}{\bm{r}}
\newcommand{\bu}{\bm{u}}
\newcommand{\bv}{\bm{v}}
\newcommand{\bw}{\bm{w}}
\newcommand{\bx}{\bm{x}}
\newcommand{\by}{\bm{y}}
\newcommand{\bz}{\bm{z}}

\newcommand{\bR}{\bm{R}}

\newcommand{\bzero}{\bm{0}}

\newcommand{\CC}{\mathbb{C}}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\ZZ}{\mathbb{Z}}

\newcommand{\eps}{\epsilon}
\newcommand{\grad}{\nabla}
\newcommand{\lam}{\lambda}
\newcommand{\lap}{\triangle}

\newcommand{\ip}[2]{\ensuremath{\left<#1,#2\right>}}

\newcommand{\image}{\operatorname{im}}
\newcommand{\onull}{\operatorname{null}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\range}{\operatorname{range}}
\newcommand{\trace}{\operatorname{tr}}

\newcommand{\ds}{\displaystyle}

\newcommand{\prob}[1]{\bigskip\noindent\textbf{#1.}\quad }
\newcommand{\probset}[2]{\bigskip\noindent\textbf{from Problem Set #1, pages #2:}\quad }

\newcommand{\pts}[1]{(\emph{#1 pts}) }
\newcommand{\epart}[1]{\medskip\noindent\textbf{(#1)}\quad }
\newcommand{\ppart}[1]{\,\textbf{(#1)}\quad }

\newcommand{\Matlab}{\textsc{Matlab}\xspace}
\newcommand{\Octave}{\textsc{Octave}\xspace}


\begin{document}
\scriptsize \noindent Math 314 Linear Algebra (Bueler) \hfill Spring 2022
\normalsize\medskip

\Large
\centerline{\textbf{Homework \#7}}

\normalsize
\bigskip
\begin{quote}
Due Monday 14 March, 2022 at 11:59pm.

\medskip
\noindent Submit as a single PDF via Gradescope, linked from the Canvas page

\href{https://canvas.alaska.edu/courses/7017}{\texttt{canvas.alaska.edu/courses/7017}}

\noindent Textbook Problems from Strang, \emph{Intro Linear Algebra}, 5th ed.~will be graded for completion.  Answers/solutions are linked at

\href{https://bueler.github.io/math314/resources.html}{\texttt{bueler.github.io/math314/resources.html}}

\noindent \textbf{P} Problems will be graded for correctness.  When grading these, I expect you to write explanations using complete sentences!
\end{quote}
\medskip

\thispagestyle{empty}

\noindent \hrulefill

\noindent \emph{Put these Textbook Problems first on your PDF, in this order.}

\probset{3.3}{157--162} \# 3, 4, 10, 16, 18, 21, 30

\probset{3.4}{174--179} \# 1, 2, 11, 13, 16, 32, 38


\bigskip
\noindent \hrulefill

\noindent \emph{Put these \textbf{P} Problems next on your PDF, in this order.}

\prob{P28}  Include $\bb$ as a fourth column in elimination on the following system, that is, use an augmented matrix.  Identify a condition on $b_1,b_2,b_3$ so that this system is solvable:
\begin{align*}
x + 2 y - 2 z &= b_1 \\
2 x + 5 y - 4 z &= b_2 \\
-2 x - 6 y + 4 z &= b_3
\end{align*}


\prob{P29}  Give examples of matrices $A$ for which the number of solutions to $A \bx = \bb$ is

\renewcommand{\labelenumi}{\textbf{(\alph{enumi})}}
\begin{enumerate}
\item 0 or 1, depending on $\bb$
\item $\infty$, regardless of $\bb$
\item 0 or $\infty$, depending on $\bb$
\item 1, regardless of $\bb$
\end{enumerate}


\prob{P30}  \ppart{a} If $A\bx = \bb$ has two distinct solutions $\bx_1$ and $\bx_2$, find two solutions of $A \bx=\bzero$.

\epart{b} Then find another (distinct) solution to $A \bx = \bb$.


\prob{P31}  Find matrices $A$ and $B$ with the given property, or explain why you cannot:

\epart{a} the only solution to $A\bx = \begin{bmatrix} 1 \\ 2 \\ 3 \end{bmatrix}$ is $\bx=\begin{bmatrix} 1 \\ 0 \end{bmatrix}$

\epart{b} the only solution to $B\bx = \begin{bmatrix} 1 \\ 0 \end{bmatrix}$ is $\bx=\begin{bmatrix} 1 \\ 2 \\ 3 \end{bmatrix}$


\prob{P32}  Suppose $\bw_1,\bw_2,\bw_3$ are linearly-independent vectors in $\RR^n$.

\epart{a}  Show that $\bv_1=\bw_2-\bw_3$, $\bv_2=\bw_1-\bw_3$, and $\bv_3=\bw_1-\bw_2$ are linearly-\emph{dependent}.

\epart{b}  Show that $\bz_1=\bw_2+\bw_3$, $\bz_2=\bw_1+\bw_3$, and $\bz_3=\bw_1+\bw_2$ are linearly-\emph{independent}.


\prob{P33}  For each part, describe \emph{all} cases where the rank is $2$.

\epart{a}  For which numbers $a$ and $b$ does the matrix $A = \begin{bmatrix}
a & b \\ b & a \end{bmatrix}$ have rank 2?

\epart{b}  For which numbers $c$ and $d$ does the matrix $B = \begin{bmatrix}
1 & 3 & 2 & 0 & 5 \\ 0 & 0 & c & 2 & 1 \\ 0 & 0 & 0 & d & 1 \end{bmatrix}$ have rank 2?


\prob{P34}  Find a basis for, and the dimension of, these subspaces of $3 \times 3$ matrices:

\epart{a} diagonal matrices

\epart{b} symmetric matrices ($A^\top = A$)

\epart{c} skew-symmetric matrices ($A^\top = -A$)


\prob{P35}  Suppose $f_1(x),f_2(x),f_3(x)$ are three functions defined on the real line.  The vector space they span could have dimension 1, 2, or 3.  Give examples $\{f_1,f_2,f_3\}$ to show each possibility.  (\emph{Each example is a list of three functions, and an explanation.})

\end{document}
